---
title: "Ch 1 excercises"
author: "Julia Grabow"
date: "2023-08-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Packages used
```{r}
library(tidyverse)
library(ggExtra)
```
Exercise 1.1:
```{r}
x1 <-
  c(3, 4, 2, 6, 8, 2, 5)
x2 <-
  c(5, 5.5, 4, 7, 10, 5, 7.5)
```

Mean of x1 and x2:

```{r}
mean(x1)

mean(x2)
```

Variances of s11 and s22 and covariance s12:
```{r}
var(x1)

var(x2)

cov(x1, x2)
```

Exercise 1.2
```{r}
age_yrs <-
  c(1, 2, 3, 4, 5, 6, 7, 8, 9, 11)

price_dollars <-
  c(18.95, 19.00, 17.95, 15.54, 14.00, 12.95, 8.94, 7.49, 6.00, 3.9)
```

a) Scatterplot with dot diagram or histograms:
```{r}
p <-
  ggplot() +
  geom_point(
    mapping = aes(
      x = age_yrs,
      y = price_dollars
    )
  )

ggMarginal(
  p,
  type = "histogram"
)
```

b) infer sign of sample covariance:
As age increases, price decreases, so I expect this to have a negative covariance

c)
Sample mean of age:
```{r}
mean(age_yrs)
```
```{r}
mean_age <- mean(age_yrs)
```

Sample mean of price_dollars:
```{r}
mean(price_dollars)
```
```{r}
mean_price <-mean(price_dollars)
```

Sample variance of age:
```{r}
var(age_yrs)
```
```{r}
var_age <- var(age_yrs)
```

Sample variance of price
```{r}
var(price_dollars)
```
```{r}
var_price <- var(price_dollars)
```

Sample covariance of age and price:
```{r}
cov(age_yrs, price_dollars)
```
```{r}
covariance_age_price <- cov(age_yrs, price_dollars)
```

Sample correlation coefficient:
```{r}
cor(age_yrs, price_dollars)
```
```{r}
correlation_age_price <- cor(age_yrs, price_dollars)
```
Sample mean array (x bar):
```{r}
array(
  list(
    mean_age,
    mean_price
  ),
  dim = c(
    2, 1
  ) 
)
```
Sample variance-covariance array (Sn):
```{r}
array(
  data = list(
    var_age,
    covariance_age_price,
    covariance_age_price,
    var_price
  ),
  dim = c(
    2, 2
  )
)
```
Sample correlation array (R):
```{r}
array(
  data = list(
    1, correlation_age_price, correlation_age_price, 1
  ),
  dim = c(2, 2)
)
```

1.3
Values:
```{r}
x3.1 <-
  c(9, 2, 6, 5, 8)
x3.2 <-
  c(12, 8, 6, 4, 10)
x3.3 <-
  c(3, 4, 0, 2, 1)
```
Sample mean array:
```{r}
array(
  data = list(
    mean(x3.1),
    mean(x3.2),
    mean(x3.3)
  ),
  dim = c(3, 1)
)
```
Sample variances and covariances:
```{r}
array(
  data = list(
    var(x3.1),
    cov(x3.1, x3.2),
    cov(x3.1, x3.3),
    cov(x3.2, x3.1),
    var(x3.2),
    cov(x3.2, x3.3),
    cov(x3.3, x3.1),
    cov(x3.3, x3.2),
    var(x3.3)
  ),
  dim = c(3, 3)
)
```
Sample correlation array:
```{r}
array(
  data = list(
    1, 
    cor(x3.1, x3.2), 
    cor(x3.1, x3.3),
    cor(x3.2, x3.1), 
    1, 
    cor(x3.2, x3.3),
    cor(x3.3, x3.1), 
    cor(x3.3, x3.2), 
    1
  ),
  dim = c(3, 3)
)
```
1.4
Create dataframe:
```{r}
company <-
  c(
    "Citigroup",
"General Electric",
"American Int! Group",
"Bank of America",
"HSBCGroup",
"ExxonMobil",
"Royal Dutch/Shell",
"BP",
"INGGroup",
"Toyota Motor"
)
sales_billions <-
  c(
    108.28,
152.36,
95.04,
65.45,
62.97,
263.99,
265.19,
285.06,
92.01,
165.68
  )
profits_billions <-
  c(
    17.05,
16.59,
10.91,
14.14,
9.52,
25.33,
18.54,
15.73,
8.10,
11.13
  )
assets_billions <-
  c(
    1484.10,
750.33,
766.42,
1110.46,
1031.29,
195.26,
193.83,
191.11,
1175.16,
211.15
  )
df4 <-
  data.frame(
    company,
    sales_billions,
    profits_billions,
    assets_billions
  )
print(df4)
```
Scatterplot with marginal histograms of sales and profits:
```{r}
p4 <-
  ggplot() +
  geom_point(
    mapping = aes(
      x = sales_billions,
      y = profits_billions,
      color = company
    )
  )
ggMarginal(
  p4,
  type = "histogram"
)
```

b)
Sample means:
```{r}
colMeans(
  df4[
    sapply(
      df4, is.numeric
    )
  ]
)
```
Variance and covariance matrix:
```{r}
cov(
  df4[
    sapply(
      df4, is.numeric
    )
  ]
)
```
```{r}

```
Correlation matrix:
```{r}
cor(
  df4[
    sapply(
      df4, is.numeric
    )
  ]
)
```

1.5
Plot scatter plots and marginal histograms
Sales and assets:
```{r}
p5a <-
  ggplot() +
  geom_point(
    data = df4,
    mapping = aes(
      x = sales_billions,
      y = assets_billions,
      color = company
    )
  )

ggMarginal(
  p5a,
  type = "histogram"
)
```
Profits and Assets:
```{r}
p5b <-
ggplot() +
  geom_point(
    data = df4,
    mapping = aes(
      x= profits_billions,
      y = assets_billions,
      color = company
    )
  )

ggMarginal(
  p5b,
  type ="histogram"
)
```

1.7
Create dataframe for data:
```{r}
obs1 <- c(2, 1)
obs2 <- c(3, 2)
obs3 <- c(4, 4)

df7 <-
  data.frame(
    obs1,
    obs2,
    obs3,
    row.names = c(
      "Variable 1",
      "Variable 2"
    )
  )

print(df7)
```
Plot pairs as scatterplot:
```{r}
pairs(df7)
```

