---
title: "Ch 1 excercises"
author: "Julia Grabow"
date: "2023-08-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages used
```{r}
library(tidyverse)
library(ggExtra)
```
Exercise 1.1:
```{r}
x1 <-
  c(3, 4, 2, 6, 8, 2, 5)
x2 <-
  c(5, 5.5, 4, 7, 10, 5, 7.5)
```

Mean of x1 and x2:

```{r}
xbar1 <- mean(x1)

xbar2 <- mean(x2)

print(xbar1)
print(xbar2)
```

Variances of s11 and s22 and covariance s12:
Remember, in this textbook, author uses population variance and covariance
rather than sample variance and covariance
s11
```{r}
(1/7) * sum(
  ((x1-xbar1)^2)
)
```
s22
```{r}
(1/7) * sum(
  ((x2-xbar2)^2)
)
```
s12
```{r}
(1/7) * sum(
  ((x1-xbar1) * (x2-xbar2))
)
```

Exercise 1.2
```{r}
age_yrs <-
  c(1, 2, 3, 4, 5, 6, 7, 8, 9, 11)

price_dollars <-
  c(18.95, 19.00, 17.95, 15.54, 14.00, 12.95, 8.94, 7.49, 6.00, 3.9)
```

a) Scatterplot with dot diagram or histograms:
```{r}
p <-
  ggplot() +
  geom_point(
    mapping = aes(
      x = age_yrs,
      y = price_dollars
    )
  )

ggMarginal(
  p,
  type = "histogram"
)
```

b) infer sign of sample covariance:
As age increases, price decreases, so I expect this to have a negative covariance

c)
Sample mean of age:
```{r}
mean_age <- mean(age_yrs)
print(mean_age)
```

Sample mean of price_dollars:
```{r}
mean_price <- mean(price_dollars)
print(mean_price)
```

Sample variance of age:
```{r}
var(age_yrs)
```
```{r}
var_age <- var(age_yrs)
```

Sample variance of price
```{r}
var(price_dollars)
```
```{r}
var_price <- var(price_dollars)
```

Sample covariance of age and price:
```{r}
cov(age_yrs, price_dollars)
```
```{r}
covariance_age_price <- cov(age_yrs, price_dollars)
```

Sample correlation coefficient:
```{r}
cor(age_yrs, price_dollars)
```
```{r}
correlation_age_price <- cor(age_yrs, price_dollars)
```
Sample mean array (x bar):
```{r}
array(
  list(
    mean_age,
    mean_price
  ),
  dim = c(
    2, 1
  ) 
)
```
Sample variance-covariance array (Sn):
```{r}
array(
  data = list(
    var_age,
    covariance_age_price,
    covariance_age_price,
    var_price
  ),
  dim = c(
    2, 2
  )
)
```
Sample correlation array (R):
```{r}
array(
  data = list(
    1, correlation_age_price, correlation_age_price, 1
  ),
  dim = c(2, 2)
)
```

1.3
Values:
```{r}
x3.1 <-
  c(9, 2, 6, 5, 8)
x3.2 <-
  c(12, 8, 6, 4, 10)
x3.3 <-
  c(3, 4, 0, 2, 1)
```
Sample mean array:
```{r}
array(
  data = list(
    mean(x3.1),
    mean(x3.2),
    mean(x3.3)
  ),
  dim = c(3, 1)
)
```
Sample variances and covariances:
```{r}
array(
  data = list(
    var(x3.1),
    cov(x3.1, x3.2),
    cov(x3.1, x3.3),
    cov(x3.2, x3.1),
    var(x3.2),
    cov(x3.2, x3.3),
    cov(x3.3, x3.1),
    cov(x3.3, x3.2),
    var(x3.3)
  ),
  dim = c(3, 3)
)
```
Sample correlation array:
```{r}
array(
  data = list(
    1, 
    cor(x3.1, x3.2), 
    cor(x3.1, x3.3),
    cor(x3.2, x3.1), 
    1, 
    cor(x3.2, x3.3),
    cor(x3.3, x3.1), 
    cor(x3.3, x3.2), 
    1
  ),
  dim = c(3, 3)
)
```
1.4
Create dataframe:
```{r}
company <-
  c(
    "Citigroup",
"General Electric",
"American Int! Group",
"Bank of America",
"HSBCGroup",
"ExxonMobil",
"Royal Dutch/Shell",
"BP",
"INGGroup",
"Toyota Motor"
)
sales_billions <-
  c(
    108.28,
152.36,
95.04,
65.45,
62.97,
263.99,
265.19,
285.06,
92.01,
165.68
  )
profits_billions <-
  c(
    17.05,
16.59,
10.91,
14.14,
9.52,
25.33,
18.54,
15.73,
8.10,
11.13
  )
assets_billions <-
  c(
    1484.10,
750.33,
766.42,
1110.46,
1031.29,
195.26,
193.83,
191.11,
1175.16,
211.15
  )
df4 <-
  data.frame(
    company,
    sales_billions,
    profits_billions,
    assets_billions
  )
print(df4)
```
Scatterplot with marginal histograms of sales and profits:
```{r}
p4 <-
  ggplot() +
  geom_point(
    mapping = aes(
      x = sales_billions,
      y = profits_billions,
      color = company
    )
  )
ggMarginal(
  p4,
  type = "histogram"
)
```

b)
Sample means:
```{r}
colMeans(
  df4[
    sapply(
      df4, is.numeric
    )
  ]
)
```
Variance and covariance matrix:
```{r}
cov(
  df4[
    sapply(
      df4, is.numeric
    )
  ]
)
```
```{r}

```
Correlation matrix:
```{r}
cor(
  df4[
    sapply(
      df4, is.numeric
    )
  ]
)
```

1.5
Plot scatter plots and marginal histograms
Sales and assets:
```{r}
p5a <-
  ggplot() +
  geom_point(
    data = df4,
    mapping = aes(
      x = sales_billions,
      y = assets_billions,
      color = company
    )
  )

ggMarginal(
  p5a,
  type = "histogram"
)
```
Profits and Assets:
```{r}
p5b <-
ggplot() +
  geom_point(
    data = df4,
    mapping = aes(
      x= profits_billions,
      y = assets_billions,
      color = company
    )
  )

ggMarginal(
  p5b,
  type ="histogram"
)
```

1.8
P = (-1, -1), Q = (1, 0), p = 2, a11 = 1/3, a12 = 1/9, a22 = 4/27
Find Euclidean distance (simple pythagorean theorem):
```{r}
first <- (-1 - 1)^2
second <- ((-1)-0)^2
sum_of_squares <- first + second
sqrt(sum_of_squares)
```
Find statistical distance:
```{r}
third <- (-1-1)*((-1)-0)
a11 <- 1/3
a22 <- 4/27
a12 <- 1/9
sqrt(
  (first * a11) * (2 * a12 * third) * (second * a22)
)
```

1.9
Create data frame:
```{r}
x1 <- c(-6, -3, -2, 1, 2, 5, 6, 8)
x2 <- c(-2, -3, 1, -1, 2, 1, 5, 3)
df9 <- data.frame(
  x1,
  x2
)
print(df9)
```
Plot scatterplot:
```{r}
ggplot() +
  geom_point(
    data = df9,
    mapping = aes(
      x= x1,
      y = x2
    )
  )
```
Compute s11, s22, and s12
variance of x1:
```{r}
var(df9 $ x1)
```
variance of x2:
```{r}
var(df9 $ x2)
```
Covariance of x1 and x2:
```{r}
cov(df9 $ x1, df9 $ x2)
```
Find new axes for correlated x1 and x2, which is tilde x1 and tilde x2:
theta = 26 degrees, cos (26) = 0.899 and sin (26) = 0.438, x1 = -6, x2 = -2
tilde x1:
```{r}
x1_tilde <-
  (
    (x1*0.899) + (x2 * 0.438)
    )

print(x1_tilde)
```
tilde x2:
```{r}
x2_tilde <-
  (
    (df9$x1 * 0.438) + (df9$x2 *0.889)
  )

print(x2_tilde)
```
Are the tilde x1 and tilde x2 new axes? Or new points? Need to clarify this
The text calls these rotated coordinates, makes me believe new points to plot
if shifted axis is moved to fit the original axis of the plot

Find sample variances using tilde x1 and tilde x2:
tilde s1:
```{r}
s1_tilde <-
  var(x1_tilde)

print(s1_tilde)
```
tilde s2:
```{r}
s2_tilde <-
  var(x2_tilde)

print(s2_tilde)
```
Using point (4, -2) transform in relation to shifted axes and calculate distance
from the origin. Adjusted x value:
```{r}
Xadjusted <-
   (4*0.899) + (-2 * 0.438)
Yadjusted <-
  (4 * 0.438) + (-2 *0.889)
print(Xadjusted)
```
Adjusted y value:
```{r}
print(Yadjusted)
```
New coordinates = (2.72,-0.026); points moved right and up
Find distance of coordinate from origin (does the origin move when axes
are adjusted?)
```{r}
sqrt(
  ((Xadjusted^2)/(s1_tilde)) +
    ((Yadjusted^2)/(s2_tilde))
)
```
I would like to practice adding an ellipse to the plot:
```{r}
ggplot(
  mapping = aes(
      x= x1_tilde,
      y = x2_tilde
    )
) +
  geom_point() +
  stat_ellipse(
    type = "norm"
  )
```


